Abstractions for Multi-Sorted Substitutions

### 2024-03-11 paper/paper.tex
Update.

@@

  In this section, we extend our framework with an abstraction that
  allows the composition of arbitrary maps.
  The core property of composition, which we call
  \ACode{\AField{⋯-fusion}}, states that applying two maps \ACode{ϕ₁}
  and \ACode{ϕ₂} in sequence to a term \ACode{t}, is the same as
  applying their composition \ACode{(ϕ₁~\AField{·ₖ}~ϕ₂)} to \ACode{t},
  i.e.
  $$
  \ACode{(t \AField{⋯} ϕ₁) \AField{⋯} ϕ₂ \ACon{≡} t \AField{⋯} (ϕ₁ \AField{·ₖ} ϕ₂)}
  $$
  This property equips our framework with the power to
  reason about the application of multiple maps by reasoning about
  the application of a single map.
  As such it forms the basis for all lemmas involving multiple maps,
  e.g.\ that applying a weakening and then a singleton
  substitution cancel each other out.

@@

  \subsection{An Examination of Composition}
  \label{sec:composition:example}
  To motivate our abstraction for composition, we first look
  at the four compositions individually:
  \ExampleCode\KExFourComps
  The definitions reveal two interesting properties:
  \begin{enumerate}
  \item
    If we compose two maps \ACode{ϕ₁} and \ACode{ϕ₂}, then the
    resulting map is a renaming, iff both \ACode{ϕ₁} and \ACode{ϕ₂} are
    renamings.
    In other words:
    if \ACode{ϕ₁} is a \ACode{K₁}-map and \ACode{ϕ₂} is a \ACode{K₂}-map,
    then the result is a \ACode{(K₁ ⊔ K₂)}-map, where \ACode{⊔} refers to the
    lattice for \ACode{\{ \ADef{Kᵣ} , \ADef{Kₛ} \}}
    generated by \ACode{\ADef{Kᵣ} < \ADef{Kₛ}}.
  \item
    All four compositions first apply \ACode{ϕ₁} to \ACode{x},
    and then apply \ACode{ϕ₂} to the result.
    If \ACode{ϕ₁} is a renaming, then this result is another variable, but
    if \ACode{ϕ₁} is a substitution, then this result is a term.
  \end{enumerate}
  With the \ACode{\ACon{Kit}} abstraction, we can easily abstract over
  \ACode{ϕ₂} being a renaming or a substitution:
  \ExampleCode\KExTwoComps

@@

  % \LibCode\KContextHelper
  \LibCode\KContexts
  To use a type returned from a typing context, it needs to be
  weakened appropriately through the \ACode{\ADef{wk-telescope}} function:
  \LibCode\KContextLookup
  We then define a variable typing as followed:
  \LibCode\KVariableTyping

>>>>

  an implementation as an Agda framework featuring
  a reflection algorithm, representation independence for
  substitutions and typing contexts, heterogeneous equality between
  renamings and substitutions, and absence of postulated axioms.
  \begin{code}
fooo
  \end{code}

@@

  In this section, we extend our framework with an abstraction that
  allows the composition of arbitrary maps.
  The core property of composition, which we call
  \ACode{\AField{⋯-fusion}}, states that applying two maps \ACode{ϕ₁}
  and \ACode{ϕ₂} in sequence to a term \ACode{t}, is the same as
  applying their composition \ACode{(ϕ₁~\AField{·ₖ}~ϕ₂)} to \ACode{t},
  i.e. \ACode{(t \AField{⋯} ϕ₁) \AField{⋯} ϕ₂ \ACon{≡} t \AField{⋯} (ϕ₁ \AField{·ₖ} ϕ₂)}.
  % $$
  % \ACode{(t \AField{⋯} ϕ₁) \AField{⋯} ϕ₂ \ACon{≡} t \AField{⋯} (ϕ₁ \AField{·ₖ} ϕ₂)}
  % $$
  This property equips our framework with the power to
  reason about the application of multiple maps by reasoning about
  the application of a single map.
  As such it forms the basis for all lemmas involving multiple maps,
  e.g.\ that applying a weakening and then a singleton
  substitution cancel each other out.

@@

  \subsection{An Examination of Composition}
  \label{sec:composition:example}
  To motivate our abstraction, we first look
  at the four compositions individually:

  \noindent
  \begin{minipage}[t]{0.60\linewidth}
  \ExampleCode\KExFourCompsI
  \end{minipage}
  \begin{minipage}[t]{0.40\linewidth}
  \ExampleCode\KExFourCompsII
  \end{minipage}

  The definitions reveal two interesting properties:
  \begin{enumerate}
  \item
    If we compose two maps \ACode{ϕ₁} and \ACode{ϕ₂}, then the
    resulting map is a renaming, iff both \ACode{ϕ₁} and \ACode{ϕ₂} are
    renamings.
    In other words:
    if \ACode{ϕ₁} is a \ACode{K₁}-map and \ACode{ϕ₂} is a \ACode{K₂}-map,
    then the result is a \ACode{(K₁ ⊔ K₂)}-map, where \ACode{⊔} refers to the
    lattice for \ACode{\{ \ADef{Kᵣ} , \ADef{Kₛ} \}}
    generated by \ACode{\ADef{Kᵣ} < \ADef{Kₛ}}.
  \item
    All four compositions first apply \ACode{ϕ₁} to \ACode{x},
    and then apply \ACode{ϕ₂} to the result.
    If \ACode{ϕ₁} is a renaming, then this result is another variable, but
    if \ACode{ϕ₁} is a substitution, then this result is a term.
  \end{enumerate}
  With the \ACode{\ACon{Kit}} abstraction, we can easily abstract over
  \ACode{ϕ₂} being a renaming or a substitution:
  \ExampleCode\KExTwoComps

@@

  % \LibCode\KContextHelper
  \noindent
  \begin{minipage}{1.0\linewidth}
    \begin{minipage}{0.5\linewidth}
      \LibCode\KContexts
      \vspace{3mm}
    \end{minipage}
    \begin{minipage}{0.5\linewidth}
      \LibCode\KContextExt
    \end{minipage}
  \end{minipage}

  To use a type returned from a typing context, it needs to be
  weakened appropriately through the \ACode{\ADef{wk-telescope}} function:

  \noindent
  \begin{minipage}{0.5\linewidth}
    \LibCode\KContextLookupI
  \end{minipage}
  \begin{minipage}{0.5\linewidth}
    \LibCode\KContextLookupII
  \end{minipage}
  % \LibCode\KContextLookup

  We then define a variable typing as followed:
  \LibCode\KVariableTyping
